Data Exploration

### 2024-02-01 analysis/documents/1-data-preparation.qmd
update all
add plot

RoF_CameraDeploymentLog_WLU <- 
  readxl::read_excel(here::here("data-raw", "tabular", "RoF_CameraDeploymentLog.xlsx")) |> 
  janitor::clean_names()

@@

map <- ggmap::get_stadiamap(bbox = c(left = -90, bottom = 50, right = -83, top = 55.5), 
                            zoom = 5, maptype = "stamen_terrain")

>>>>

RoF_CameraDeploymentLog_WLU <- 
  suppressWarnings(readxl::read_excel(here::here("data-raw", "tabular", "RoF_CameraDeploymentLog.xlsx"),
                                      guess_max = 10)) |>
  janitor::clean_names() |> 
  dplyr::mutate(date_retrieved = as.Date.character(date_retrieved))

@@

map <- ggmap::get_stadiamap(bbox = c(left = -90, bottom = 50, right = -83, top = 55.5), 
                            zoom = 3, maptype = "stamen_terrain")

### 2024-02-01 analysis/documents/2-data-exploration.qmd
update
add plot

In this exploration of the data, we will make use of the code provided in "[Introduction to Camera Trap Data Management and Analysis in R](https://wildcolab.github.io/Introduction-to-Camera-Trap-Data-Management-and-Analysis-in-R/index.html)", hereafter refered to as ICTDMA. I adapt the code in this document into functions which should later serve as a basis for a R package.

@@

```{r, eval=FALSE}
 [1] "project_id"                          "project_name"                        "project_short_name"                 
 [4] "project_objectives"                  "project_species"                     "project_species_individual"         
 [7] "project_sensor_layout"               "project_sensor_layout_targeted_type" "project_bait_use"                   
[10] "project_bait_type"                   "project_stratification"              "project_stratification_type"        
[13] "project_sensor_method"               "project_individual_animals"          "project_blank_images"               
[16] "project_sensor_cluster"              "project_admin"                       "project_admin_email"                
[19] "project_admin_organization"          "country_code"                        "embargo"                            
[22] "initiative_id"                       "metadata_license"                    "image_license"                      
[25] "data_citation"                       "count_optional"                      "project_type
```

@@

```{r}
image_data <- readRDS(here::here("data", "objects", "annotations_animals.rds")) |> 
  dplyr::select(id, source_file_with_dep,
                image_id, species:comment, camera_id,
                maker_notes_sequence:maker_notes_user_label, plot_id:ymd) |> 
  dplyr::rename(file_path = source_file_with_dep)
head(image_data)
```

And then reformat and rename:

#### Deployment data

This is

#### Camera data

>>>>

In this exploration of the data, we will loosely follow the protocol in "[Introduction to Camera Trap Data Management and Analysis in R](https://wildcolab.github.io/Introduction-to-Camera-Trap-Data-Management-and-Analysis-in-R/index.html)", hereafter referred to as ICTDMA.

@@

```{r, eval=FALSE}
#| label: project
 [1] "project_id"                          "project_name"                        "project_short_name"                 
 [4] "project_objectives"                  "project_species"                     "project_species_individual"         
 [7] "project_sensor_layout"               "project_sensor_layout_targeted_type" "project_bait_use"                   
[10] "project_bait_type"                   "project_stratification"              "project_stratification_type"        
[13] "project_sensor_method"               "project_individual_animals"          "project_blank_images"               
[16] "project_sensor_cluster"              "project_admin"                       "project_admin_email"                
[19] "project_admin_organization"          "country_code"                        "embargo"                            
[22] "initiative_id"                       "metadata_license"                    "image_license"                      
[25] "data_citation"                       "count_optional"                      "project_type
```

@@

```{r}
#| label: image
image_data <- readRDS(here::here("data", "objects", "annotations_animals.rds")) |> 
  dplyr::select(id, source_file_with_dep,
                image_id, species:comment, camera_id,
                maker_notes_sequence:maker_notes_user_label, plot_id:ymd) |> 
  dplyr::rename(file_path = source_file_with_dep)
head(image_data)
```

#### Camera and deployment data

The camera and deployment data are currently together as each camera was only deployed once

```{r}
#| label: camdep
cam_dep_data <- readRDS(here::here("data", "objects", "cams_sf_coords.rds")) |> 
  sf::st_drop_geometry() |> 
  dplyr::select(dplyr::ends_with("_id"), date_deployed, date_retrieved, longitude, latitude)
stopifnot(sum(table(cam_dep_data$camera_id) != 1) == 0)
```

#### Deployment plot

```{r, fig.height=20, fig.width=8}
#| label: camdep-plot
cam_dep_data |> 
  dplyr::mutate(date_deployed = as.Date(date_deployed),
                date_retrieved = as.Date(date_retrieved)) |> 
  dplyr::mutate(status = ifelse(date_deployed < date_retrieved, "valid", "invalid")) |> 
  dplyr::mutate(status = ifelse(is.na(status), "invalid", status)) |> 
  ggplot2::ggplot() +
  ggplot2::geom_segment(ggplot2::aes(x = date_deployed, y = camera_id, 
                                     xend = date_retrieved, yend = camera_id,
                                     color = status)) +
  ggplot2::geom_point(ggplot2::aes(x = date_deployed, y = camera_id, 
                                     color = status)) + 
    ggplot2::geom_point(ggplot2::aes(x = date_retrieved, y = camera_id, 
                                     color = status)) + 
  ggplot2::scale_color_discrete(type = c("#E69F00","#56B4E9")) +
  ggplot2::theme_bw()
```
